Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-15 03-time.Rmd
More edits

## Reshape data for models

### Create predictors

We then create predictors (centered, lagged, rescaled) for models in the ITU dataset.

```{r}
# For models, year is centred at 
itu <- itu %>% 
  mutate(time = (year - 2013) / 10)

# Create predictor: Lagged effect of 10% increase in internet
itu <- itu %>% 
  arrange(country, year) %>% 
  group_by(country) %>% 
  mutate(i1 = lag(internet*10, 1)) %>% 
  ungroup()

# Centering within and between countries
itu %>%
  bmlm::isolate("i1", by = "country", which = "both")

tibble(
  id = c(1,1,1,2,2,2,3,3,3),
  x = c(1,2,3,4,5,6,7,8,9)
) %>% 
  bmlm::isolate("x", by = "id", which = "both") %>% 
  group_by(id) %>% 
  mutate(x_cb2 = mean(x), x_cw2 = x - x_cb2) %>% 
  ungroup() %>% 
  mutate(x_cb2 = x_cb2 - mean(x))
```

@@

@@

### GBD

#### brms

>>>>

## Prepare data for models

We prepare the predictors
- center year on 2019 (last year in outcome data)
- create 1-year lagged internet adoption variables

```{r}
# For models, year is centred at 2019
itu <- itu %>% 
  mutate(year = year - 2019)

# Create lagged predictors
itu <- itu %>% 
  arrange(country, year) %>% 
  group_by(country) %>% 
  mutate(
    across(
      c(internet, fixed, mobile), 
      list(`1` = ~lag(., 1)), 
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  )

# Centering within and between countries
itu <- itu %>%
  group_by(country) %>% 
  mutate(
    i1_cb = mean(i1_cb), # Country average
    i1_cw = i1 - i1_cb # Year-specific deviation from country average
  ) %>% 
  ungroup() %>% 
  # Grand-mean center the country averages
  mutate(i1_cb = i1_cb - mean(i1))
```

@@

Merge tables

```{r}
dat <- left_join(gbd, itu)
```

@@

## GBD

### brms
